---
title: "HW04"
author: "Elliot Smith"
date: "4/2/2018"
output: pdf_document
---

```{r, echo = FALSE}

## Load in the necessary libraries

suppressMessages(
    library(lme4)
)

```

# Problem 1

```{r, echo = FALSE}

## Load in the data

data_1 <- read.csv("~/Documents/Rice_University/Spring_2018/STAT616/HW04/blood.csv")

```

## Part a

### ANOVA Table

```{r, echo = FALSE}

data_1a <- setNames(data.frame(matrix(NA, ncol = 2, nrow = 45)), c("doctor", "value"))
data_1a$doctor <- c(rep("doc1", times = 15), rep("doc2", times = 15), rep("doc3", times = 15))
data_1a$value <- c(data_1$doc1, data_1$doc2, data_1$doc3)

mu_doc <- mean(data_1a$value)

lm_1a <- lm(value ~ doctor, data = data_1a)
# lm_1a <- lm(value ~ doctor, data = data_1a, contrasts = list(doctor = contr.sum))
anova(lm_1a)
# summary(lm_1a)

sigma_alpha_1a <- (248.163 - 1.784) / 15
sigma_epsilon_1a <- 1.784
# For standard error, check against results from model fit by contra-sum
se_mu_hat_1a <- sqrt(sigma_epsilon_1a / 45)
# For confidence interval, check against results from model fit by contra-sum
ci_mu_hat_1a <- confint(lm_1a)
# c(mu_doc - qt(0.975, df = 42) * se_mu_hat_1a, mu_doc + qt(0.975, df = 42) * se_mu_hat_1a)

```

### Estimates

* $\sigma_\epsilon^2 = 1.784$
* $\sigma_\alpha^2 = \frac{248.163 - 1.784}{15} = 16.42527$
* $SE(\hat\mu) = \sqrt{\frac{\sigma_\epsilon^2}{45}} = 0.1991091$
* Confidence Interval = $[129.6195, 130.4232]$

## Part b

### LM Summary

```{r, echo = FALSE}

lm_1b <- lmer(value ~ (1|doctor), data = data_1a, REML = FALSE)
summary(lm_1b)

sigma_alpha_1b <- 10.911
sigma_epsilon_1b <- 1.784
se_mu_hat_1b <- sqrt((sigma_epsilon_1b + (15 * sigma_alpha_1b)) / 45)
ci_mu_hat_1b <- confint(lm_1b)

```

### Estimates

* $\sigma_\epsilon^2 = 1.784$
* $\sigma_\alpha^2 = 10.911$
* $SE(\hat\mu) = \sqrt{\frac{\sigma_\epsilon^2 + 15\sigma_\alpha^2}{45}} = 1.917458$
* Confidence Interval = $[124.668177, 135.374484]$

## Part c

### LM Summary

```{r, echo = FALSE}

lm_1c <- lmer(value ~ (1|doctor), data = data_1a, REML = TRUE)
summary(lm_1c)

sigma_alpha_1c <- 16.425
sigma_epsilon_1c <- 1.784
se_mu_hat_1c <- sqrt((sigma_epsilon_1c + (15 * sigma_alpha_1c)) / 45)
ci_mu_hat_1c <- confint(lm_1c)

```

### Estimates

* $\sigma_\epsilon^2 = 1.784$
* $\sigma_\alpha^2 = 16.425$
* $SE(\hat\mu) = \sqrt{\frac{\sigma_\epsilon^2 + 15\sigma_\alpha^2}{45}} = 2.348328$
* Confidence Interval = $[124.668177, 135.374484]$

## Part d

* All of the methods have the same $\sigma_\epsilon^2$
* ANOVA and REML have the same $\sigma_\alpha^2$ of 16.425, while the ML method's is 10.911
* ANOVA has smallest $SE(\hat\mu)$ at 0.1991, ML is next smallest at 1.917 and REML is the largest at 2.348
* The Confidence Interval for ANOVA is the tightest by far, where REML and ML are the same and much wider
* 1a and 1c are biased estimators, 1b estimators are unbiases
* I prefer 1b (the REML method) because the estimators are unbiased

## Part e - ANOVA Method

### ANOVA Table

```{r, echo = FALSE}

data_1e <- setNames(data.frame(matrix(NA, ncol = 2, nrow = 45)), c("device", "value"))
data_1e$device <- c(rep("dev1", times = 15), rep("dev2", times = 15), rep("dev3", times = 15))
data_1e$value <- c(data_1$dev1, data_1$dev2, data_1$dev3)

mu_dev <- mean(data_1e$value)

## ANOVA

lm_1e_anova <- lm(value ~ device, data = data_1e)
# lm_1e_anova <- lm(value ~ device, data = data_1e, contrasts = list(device = contr.sum))
anova(lm_1e_anova)
# summary(lm_1e_anova)

sigma_alpha_1e_anova <- (0.007 - 88.082) / 15
sigma_alpha_1e_anova <- 0
sigma_epsilon_1e_anova <- 88.082
# For standard error, check against results from model fit by contra-sum
se_mu_hat_1e_anova <- sqrt((sigma_epsilon_1e_anova + (15 * sigma_alpha_1e_anova)) / 45)
# For confidence interval, check against results from model fit by contra-sum
ci_mu_hat_1e_anova <- confint(lm_1e_anova)
# c(mu_dev - qt(0.975, df = 42) * se_mu_hat_1e_anova, mu_dev + qt(0.975, df = 42) * se_mu_hat_1e_anova)

```

### Estimates

* $\sigma_\epsilon^2 = 88.082$
* $\sigma_\alpha^2 = \frac{0.007 - 88.082}{15} = -5.871667 = 0$
* $SE(\hat\mu) = \sqrt{\frac{\sigma_\epsilon^2}{45}} = 1.399063$
* Confidence Interval = $[125.243023, 130.889866]$

## Part e - LM Method

### LM Summary

```{r, echo = FALSE}

## ML

lm_1e_ml <- lmer(value ~ (1|device), data = data_1e, REML = FALSE)
summary(lm_1e_ml)

sigma_alpha_1e_ml <- 0
sigma_epsilon_1e_ml <- 82.21
se_mu_hat_1e_ml <- sqrt((sigma_epsilon_1e_ml + (15 * sigma_alpha_1e_ml)) / 45)
ci_mu_hat_1e_ml <- confint(lm_1e_ml)

```

### Estimates

* $\sigma_\epsilon^2 = 82.21$
* $\sigma_\alpha^2 = 0$
* $SE(\hat\mu) = \sqrt{\frac{\sigma_\epsilon^2 + 15\sigma_\alpha^2}{45}} = 1.351625$
* Confidence Interval = $[125.359749, 130.773140]$

## Part e - REML Method

### LM Summary

```{r, echo = FALSE}

## REML

lm_1e_reml <- lmer(value ~ (1|device), data = data_1e, REML = TRUE)
summary(lm_1e_reml)

sigma_alpha_1e_reml <- 0
sigma_epsilon_1e_reml <- 84.08
se_mu_hat_1e_reml <- sqrt((sigma_epsilon_1e_reml + (15 * sigma_alpha_1e_reml)) / 45)
ci_mu_hat_1e_reml <- confint(lm_1e_reml)

```

### Estimates

* $\sigma_\epsilon^2 = 84.08$
* $\sigma_\alpha^2 = 0$
* $SE(\hat\mu) = \sqrt{\frac{\sigma_\epsilon^2 + 15\sigma_\alpha^2}{45}} = 1.366911$
* Confidence Interval = $[125.359749, 130.773140]$

## Part e - Results Comparison

* The largest $\sigma_\epsilon^2$ is from ANOVA with 88.082, followed by REML with 84.08 and finally ML with 82.21
* ANOVA, ML and REML all have a $\sigma_\alpha^2$ of 0
* ML has smallest $SE(\hat\mu)$ at 1.352, REML is next smallest at 1.367 and ANOVA is the largest at 1.399
* All of the Confidence Intervals are about equally as tight, where REML and ML are the same again
* ANOVA and ML are biased estimators, REML estimators are unbiased
* I prefer the REML method because the estimators are unbiased

# Problem 2








